---
title: "MXB344_NN_version"
author: "Joe Grech"
date: "07/08/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import the data
``` {r import data }



final <- read.csv("final_data.csv")



#remove useless variables
final<-final[-c(1,2,3,4)]
final$defaulted = is.na(final$mths_since_last_delinq)
final<-final[complete.cases(final[,c("loan_amnt","funded_amnt","term","int_rate","installment","grade","sub_grade","emp_length","home_ownership","annual_inc","defaulted","verification_status","zip_code","addr_state","dti","delinq_2yrs","earliest_cr_line","inq_last_6mths","open_acc","pub_rec","revol_bal","revol_util","repay_fail")]),]



##Remove the month since last delinque
#remove last payment, next payment, last credit pull


final<-final[,c("loan_amnt","funded_amnt","term","int_rate","installment","grade","sub_grade","emp_length","home_ownership","annual_inc","defaulted","verification_status","zip_code","addr_state","dti","delinq_2yrs","earliest_cr_line","inq_last_6mths","open_acc","pub_rec","revol_bal","revol_util","repay_fail")]
colSums(is.na(final))


final<-final[ , colSums(is.na(final)) == 0]

